SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-09-18 scipy-1.0/paper.tex
Fix typos reported by Tyler

With a rich programming environment and a numerical array object in
place, the time was ripe for the development of a full scientific
software stack.  Many of the first efforts were driven by graduate
students and post-doctoral researchers, but soon attracted the
attention of senior researchers.
% Paul Dubois, CDAT
% https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
Paul Dubois at Lawrence Livermore National Laboratories took over the
maintenance of Numeric and funded the writing of its
manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}.
%in the Program for Climate Model Diagnosis and Intercomparison at
%Lawrence Livermore National Laboratory built their Climate Data
%Analysis Tools (``an open-source set of Python modules and tools for
%accessing, manipulating, and plotting scientific data in general and
%climate data sets in
%particular''\footnote{https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html})
%on Numeric, the newly established Python array object.  Paul later
%became the maintainer for Numeric, and helped fund the writing of the
%first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
In 1998, the Space Telescope Science Institute, in charge of science
operations of the Hubble Space Telescope, decided to replace their
custom scripting language with Python\footnote{https://conference.scipy.org/scipy2011/slides/greenfield\_keynote\_astronomy.pdf}.

@@

In 2000, STScI ported their entire IDL-based analysis pipeline, IRAF,
to Python.  And it was here that a small crisis developed around the
Python numerical array container.  Numeric, the original array package, was
suitable for small arrays, but not for the large images processed by
STScI.  With the Numeric maintainer's blessing, the decision was made
to write NumArray, a library that could handle data on a different
scale.  Unfortunately, NumArray proved inefficient for small arrays,
facing the community with a rather unfortunate choice.  In 2005,
Travis Oliphant combined the best elements of Numeric and NumArray,
thereby solving the dilemma and paving the way for Python to become
a very significant player in the Data Science movement. \fixme{mention
  first NumPy meeting at Berkeley}

@@

\begin{figure}
\begin{verbatim}
    2001: first SciPy release
    2005: transition to NumPy
    2007: creation of scikits
    2008: spatial module and first Cython code added
    2010: move to a 6-monthly release cycle
    2011: development moves to GitHub
    2011: Python 3 support
    2012: add a sparse graph module and unified optimization interface
    2012: remove of maxentropy
    2013: continuous integration with TravisCI
    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
    2017: add a unified C API with LowLevelCallable; remove weave
    2017: SciPy 1.0 release
\end{verbatim}
\caption{Timeline}
\end{figure}

>>>>

With a rich programming environment and a numerical array object in
place, the time was ripe for the development of a full scientific
software stack.  Many of the first efforts were driven by graduate
students and post-doctoral researchers, but soon attracted the
attention of senior researchers.
% Paul Dubois, CDAT
% https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html
Paul Dubois at Lawrence Livermore National Laboratory took over the
maintenance of Numeric and funded the writing of its
manual\footnote{http://www.numpy.org/\_downloads/numeric-manual.pdf}.
%in the Program for Climate Model Diagnosis and Intercomparison at
%Lawrence Livermore National Laboratory built their Climate Data
%Analysis Tools (``an open-source set of Python modules and tools for
%accessing, manipulating, and plotting scientific data in general and
%climate data sets in
%particular''\footnote{https://mail.python.org/pipermail/numpy-discussion/2001-May/000645.html})
%on Numeric, the newly established Python array object.  Paul later
%became the maintainer for Numeric, and helped fund the writing of the
%first manual\footnote{https://conference.scipy.org/scipy2010/slides/travis_oliphant_keynote.pdf}
In 1998, the Space Telescope Science Institute, in charge of science
operations of the Hubble Space Telescope, decided to replace their
custom scripting language with Python\footnote{https://conference.scipy.org/scipy2011/slides/greenfield\_keynote\_astronomy.pdf}.

@@

In 2000, STScI ported their entire IDL-based analysis pipeline, IRAF,
to Python.  And it was here that a small crisis developed around the
Python numerical array container.  Numeric, the original array package, was
suitable for small arrays, but not for the large images processed by
STScI.  With the Numeric maintainer's blessing, the decision was made
to write NumArray, a library that could handle data on a different
scale.  Unfortunately, NumArray proved inefficient for small arrays,
presenting the community with a rather unfortunate choice.  In 2005,
Travis Oliphant combined the best elements of Numeric and NumArray,
thereby solving the dilemma and paving the way for Python to become
a very significant player in the Data Science movement. \fixme{mention
  first NumPy meeting at Berkeley}

@@

\begin{figure}
\begin{verbatim}
    2001: first SciPy release
    2005: transition to NumPy
    2007: creation of scikits
    2008: spatial module and first Cython code added
    2010: move to a 6-monthly release cycle
    2011: development moves to GitHub
    2011: Python 3 support
    2012: add a sparse graph module and unified optimization interface
    2012: removal of maxentropy
    2013: continuous integration with TravisCI
    2015: adding Cython interface for BLAS/LAPACK and a benchmark suite
    2017: add a unified C API with LowLevelCallable; remove weave
    2017: SciPy 1.0 release
\end{verbatim}
\caption{Timeline}
\end{figure}
